Landfill Geographic Analysis: Owner Level

### 2024-08-01 trash/landfill/data_clean_owner.Rmd
added code written by Romeo Ignacio to geocode landfills.

---
title: "Landfill Geographic Analysis: Operator Level"
output: html_document
date: "2024-01-11"
author: Charles Taragin and Romeo Ignacio
---

@@

#Bestway Disposal (Different owner???)
#Monadnock Disposal Services (Diferent owner???)
#Texas Disposal Systems vs Texas Disposal Systems Inc.
operator_names <- raw %>%
  filter(Type=="LF"|Family=="LF"|Type=="TS"|Family=="TS") %>%
  select(`Operator Entity`) %>%
  unique()
#Willimantic Waste Paper Co For operator (Different operator???)
#Bestway Disposal (Different operator???)
# 9499 observations before
disposal_df <- raw %>%
  select(Code,Location,City,County,State,Zip,Longitude,Latitude,Code,Type,Family,`Waste Shed`,`Operator Code`,`Total of All Wastes (tons/day)`,Operation) %>% # Make changes to entity type here
  mutate(st_cnty=paste(County,State,sep = ":")) %>%
  rename(Waste_Shed = `Waste Shed`,
         Operator_Code = `Operator Code`)

@@

landfill_csa_df <- intersect_df %>%
  st_drop_geometry() %>%
  filter(Type=="LF"|Family=="LF"|Type=="TS"|Family=="TS") %>% # Step 1: Keep only LF basde on either Type or Family
  filter(Waste_Shed!="Exclusive or Captive"|is.na(Waste_Shed)) %>% # Step 2: Drop all observations labelled "Exclusive or Captive" (1619 Observations)
  mutate(firm_name="") %>%
  select(Operator_Code,market,st_cnty,County,State,`Total of All Wastes (tons/day)`,iscsa) %>% # Change this line
  group_by(Operator_Code,market,st_cnty,County,State,`Total of All Wastes (tons/day)`,iscsa) %>% # Change this line
  summarise(`Total of All Wastes (tons/day)`=sum(`Total of All Wastes (tons/day)`))

# Inquire about wasteshed and what to do if NA???
# Old observation count 7193
# New observation number 6823
landfill_csa_df <- intersect_df %>%
  st_drop_geometry() %>%
  filter(Type=="LF"|Family=="LF"|Type=="TS"|Family=="TS") %>% # Step 1: Keep only LF basde on either Type or Family
  filter(Waste_Shed!="Exclusive or Captive"|is.na(Waste_Shed)) %>% # Step 2: Drop all observations labelled "Exclusive or Captive" (1619 Observations)
  mutate(firm_name="") %>%
  select(Operator_Code,market,st_cnty,County,State,`Total of All Wastes (tons/day)`,iscsa,Operation) %>% # Change this line
  ungroup()

# Old observation count 1703
landfill_cbsa_df <- cbsa_intersect_df %>%
  st_drop_geometry() %>%
  filter(Type=="LF"|Family=="LF") %>% # Step 1: Keep only LF based on either Type or Family
  filter(Waste_Shed!="Exclusive or Captive"|is.na(Waste_Shed)) %>% # Step 2: Drop all observations labelled "Exclusive or Captive" (1619 Observations)
  select(Operator_Code,market,st_cnty,County,State,`Total of All Wastes (tons/day)`,iscbsa) %>% # Change this line
  ungroup()

@@

This analysis is done on the Operator Code level or Operator Entity name if Operator Code does not exist.  Let's first look at number of markets that are in a CSA and number of markets that are county based as they do not fall in a CSA.  We see that only `r paste0(round(sum_landfill$per_count[1],0),"%")` of landfills are CSA market based
```{r}

@@

```
```{r}
## Create summary statistics (On MSA-Owner-Operator Lvl)
landfill_sum_csa <- landfill_csa_df %>%
  ungroup() %>%
  group_by(market,Operator_Code) %>%
  summarise(count=n(),
            total_waste=sum(`Total of All Wastes (tons/day)`), .groups="keep")

@@

```{r}
## Create HHI Level data
### On CSA Level ###
landfill_shrs_csa <- landfill_csa_df %>%
  ungroup() %>%
  group_by(Operator_Code,market,iscsa) %>% # Aggregate on Market level, Owner Entity, Operator level
  summarise(waste_own = sum(`Total of All Wastes (tons/day)`), .groups="keep") %>%
  ungroup() %>%
  group_by(market) %>% # Aggregate on Market
  mutate(total_waste = sum(waste_own),
         waste_shr = 100*(waste_own/total_waste),
         waste_hhi = sum(waste_shr^2))
### On CBSA Level ###
landfill_shrs_cbsa <- landfill_cbsa_df %>%
  ungroup() %>%
  group_by(Operator_Code,market,iscbsa) %>% # Aggregate on Market level, Owner Entity, Operator level
  summarise(waste_own = sum(`Total of All Wastes (tons/day)`), .groups="keep") %>%
  ungroup() %>%
  group_by(market) %>% # Aggregate on Market
  mutate(total_waste = sum(waste_own),
         waste_shr = 100*(waste_own/total_waste),
         waste_hhi = sum(waste_shr^2)) ## Have NA remove option

@@

```{r}
### Custom summarise function command

@@

Now let's look at the tabulation of number of firms per market (assuming equally sized firms) from the HHIs we calculated before.  We will use the ceiling function to round to the nearest integer.

```{r}
# summary(non_mono_and_mono_landfill_csa_df$waste_hhi)
non_mono_and_mono_landfill_csa_df <- landfill_csa %>% 
  mutate(eff_firms = 1e4/waste_hhi,
         old_eff_firms=eff_firms,
         old_eff_firms=ceiling(eff_firms),
         eff_firms = ifelse((ceiling(eff_firms)-eff_firms)>0.5,floor(eff_firms),ceiling(eff_firms)))
#1.8 median means there are 1.8 equally sized firms in the market (equivalent)
```

```{r}
data.frame(table(non_mono_and_mono_landfill_csa_df$old_eff_firms)) %>%
  pivot_wider(names_from=Var1, values_from=Freq) %>% 
  kbl(full_wdith=F,
      digits=0,
      caption = "Tabulations if Firms were equally sized (Based on HHIs)") %>%
  kable_classic_2(full_width = F)
```

Now let's do a similar exericse but correcting for rounding issues.

```{r}
data.frame(table(non_mono_and_mono_landfill_csa_df$eff_firms)) %>%
  pivot_wider(names_from=Var1, values_from=Freq) %>% 
  kbl(full_wdith=F,
      digits=0,
      caption = "Tabulations if Firms were equally sized (Based on HHIs. Corrected for rounding)") %>%
  kable_classic_2(full_width = F)
```

@@

## Data check if all percentages in a market equal 1
chk_lf_df <- landfill_shrs_csa %>% 
  group_by(market) %>%
  mutate(chk_per=sum(waste_shr)) %>%
  ungroup() %>%
  select(chk_per) %>%
  unique()
```

Now let's look at the tabulation of number of firms per market (assuming equally sized firms) from the HHIs we calculated before like we did in the previous section.  Notice that the tabulations are quite similar when we looked at all markets.  The first table calculates this based on a ceiling function while the second one calculates this with some correction for rounding (using either a floor or ceiling function dependent on the number)

```{r}
# summary(non_mono_landfill_csa_df$waste_hhi)
non_mono_landfill_csa_df <- non_mono_landfill_csa_df %>% 
  mutate(eff_firms = 1e4/waste_hhi,
         old_eff_firms=eff_firms,
         old_eff_firms=ceiling(eff_firms),
         eff_firms = ifelse((ceiling(eff_firms)-eff_firms)>0.5,floor(eff_firms),ceiling(eff_firms)))
#1.8 median means there are 1.8 equally sized firms in the market (equivalent)
```

```{r}
data.frame(table(non_mono_landfill_csa_df$old_eff_firms)) %>%
  pivot_wider(names_from=Var1, values_from=Freq) %>% 
  kbl(full_wdith=F,
      digits=0,
      caption = "Tabulations if Firms were equally sized (Based on HHIs)") %>%
  kable_classic_2(full_width = F)
```

```{r}
data.frame(table(non_mono_landfill_csa_df$eff_firms)) %>%
  pivot_wider(names_from=Var1, values_from=Freq) %>% 
  kbl(full_wdith=F,
      digits=0,
      caption = "Tabulations if Firms were equally sized (Based on HHIs. Corrected for rounding)") %>%
  kable_classic_2(full_width = F)
```

### Chatanooga County

```{r}
chat_sum <- chat_shrs %>%
  ungroup() %>%
  select(Operator_Code,waste_own,waste_shr,waste_hhi)

@@

```{r}
chat_sum %>%
  select(Operator_Code,waste_own,waste_shr) %>%
  arrange(desc(waste_shr)) %>%
  kbl(col.names=c("Operator Code","Total Waste","Waste Share"),
      digits=0,
      caption = "# Landfills by Markets") %>%
  kable_classic_2(full_width = F)
```

@@

## CBSA level bar chart data
# bar_cbsa <- landfill_shrs_cbsa %>% 
#   ungroup() %>%
#   group_by(market) %>%
#   summarise(landfills_cnt = n()) %>%
#   ungroup() %>%
#   group_by(landfills_cnt) %>%
#   summarise(freq=n()) %>%
#   mutate(grp_lbls = ifelse(landfills_cnt>=10,"10+",landfills_cnt)) %>%
#   ungroup() %>%
#   group_by(grp_lbls) %>%
#   summarise(freq=sum(freq)) %>%
#   mutate(lvl="CBSA")

@@

```{r}
### Check to see if they are the same
chk_vertical <- disposal_df %>%
  filter(Type=="LF"|Family=="LF"|Type=="TS"|Family=="TS") %>%
  mutate(is_municp=grepl("COUNTY|CITY|TOWN|VILLAGE",Operator_Code),
         is_municp_pub=ifelse(Operation=="Public",TRUE,FALSE),
         is_equal=is_municp==is_municp_pub)

landfill_shrs_csa_operation <- landfill_csa_df %>%
  ungroup() %>%
  group_by(Operator_Code,market,iscsa,Operation) %>% # Aggregate on Market level, Owner Entity, Operator level
  summarise(waste_own = sum(`Total of All Wastes (tons/day)`), .groups="keep") %>%
  ungroup() %>%
  group_by(market) %>% # Aggregate on Market
  mutate(total_waste = sum(waste_own),
         waste_shr = 100*(waste_own/total_waste),
         waste_hhi = sum(waste_shr^2))
```

@@

vertical_firms <- landfill_shrs_csa %>%
  mutate(is_municp=grepl("COUNTY|CITY|TOWN|VILLAGE",Operator_Code)) %>%
  group_by(is_municp,market) %>%
  summarise(freq=n()) %>%
  mutate(is_municp= factor(is_municp,levels=c(TRUE,FALSE),labels=c("Vertically Integrated","Non-Vertically Integrated"))) %>%
  pivot_wider(names_from=is_municp, values_from=freq) %>%
  mutate(`Non-Vertically Integrated`=ifelse(is.na(`Non-Vertically Integrated`),0,`Non-Vertically Integrated`),
         `Vertically Integrated`=ifelse(is.na(`Vertically Integrated`),0,`Vertically Integrated`),
         `Non-Vertically Integrated`=factor(ifelse(`Non-Vertically Integrated`>=10,"10+",`Non-Vertically Integrated`),levels=c(seq(0,9),"10+"),labels=paste(c(seq(0,9),"10+"))),
         `Vertically Integrated`=factor(ifelse(`Vertically Integrated`>=10,"10+",`Vertically Integrated`),levels=c(seq(0,9),"10+"),labels=paste(c(seq(0,9),"10+"))))
```

>>>>

---
title: "Landfill Geographic Analysis: Owner Level"
output: html_document
date: "2024-01-11"
author: Charles Taragin and Romeo Ignacio
---

@@

#Bestway Disposal (Different owner???)
#Monadnock Disposal Services (Diferent owner???)
#Texas Disposal Systems vs Texas Disposal Systems Inc.
operator_names <- raw %>%
  filter(Type=="LF"|Family=="LF"|Type=="TS"|Family=="TS") %>%
  select(`Operator Entity`) %>%
  unique()
#Willimantic Waste Paper Co For operator (Different operator???)
#Bestway Disposal (Different operator???)
# 9499 observations before
disposal_df <- raw %>%
  select(Code,Location,City,County,State,Zip,Longitude,Latitude,Code,Type,Family,`Waste Shed`,`Owner Code`,`Total of All Wastes (tons/day)`,Operation) %>% # Make changes to entity type here
  mutate(st_cnty=paste(County,State,sep = ":")) %>%
  rename(Waste_Shed = `Waste Shed`,
         Owner_Code = `Owner Code`)

@@

landfill_csa_df <- intersect_df %>%
  st_drop_geometry() %>%
  filter(Type=="LF"|Family=="LF"|Type=="TS"|Family=="TS") %>% # Step 1: Keep only LF basde on either Type or Family
  filter(Waste_Shed!="Exclusive or Captive"|is.na(Waste_Shed)) %>% # Step 2: Drop all observations labelled "Exclusive or Captive" (1619 Observations)
  mutate(firm_name="") %>%
  select(Owner_Code,market,st_cnty,County,State,`Total of All Wastes (tons/day)`,iscsa) %>% # Change this line
  group_by(Owner_Code,market,st_cnty,County,State,`Total of All Wastes (tons/day)`,iscsa) %>% # Change this line
  summarise(`Total of All Wastes (tons/day)`=sum(`Total of All Wastes (tons/day)`))

# Inquire about wasteshed and what to do if NA???
# Old observation count 7193
# New observation number 6823
landfill_csa_df <- intersect_df %>%
  st_drop_geometry() %>%
  filter(Type=="LF"|Family=="LF"|Type=="TS"|Family=="TS") %>% # Step 1: Keep only LF basde on either Type or Family
  filter(Waste_Shed!="Exclusive or Captive"|is.na(Waste_Shed)) %>% # Step 2: Drop all observations labelled "Exclusive or Captive" (1619 Observations)
  mutate(firm_name="") %>%
  select(Owner_Code,market,st_cnty,County,State,`Total of All Wastes (tons/day)`,iscsa,Operation) %>% # Change this line
  ungroup()

# Old observation count 1703
landfill_cbsa_df <- cbsa_intersect_df %>%
  st_drop_geometry() %>%
  filter(Type=="LF"|Family=="LF") %>% # Step 1: Keep only LF based on either Type or Family
  filter(Waste_Shed!="Exclusive or Captive"|is.na(Waste_Shed)) %>% # Step 2: Drop all observations labelled "Exclusive or Captive" (1619 Observations)
  select(Owner_Code,market,st_cnty,County,State,`Total of All Wastes (tons/day)`,iscbsa) %>% # Change this line
  ungroup()

@@

This analysis is done on the Owner Code level or Owner Entity name if Owner Code does not exist.  Let's first look at number of markets that are in a CSA and number of markets that are county based as they do not fall in a CSA.  We see that only `r paste0(round(sum_landfill$per_count[1],0),"%")` of landfills are CSA market based
```{r}

@@

```
```{r}
## Create summary statistics (On MSA-Owner-Operator Lvl)
landfill_sum_csa <- landfill_csa_df %>%
  ungroup() %>%
  group_by(market,Owner_Code) %>%
  summarise(count=n(),
            total_waste=sum(`Total of All Wastes (tons/day)`), .groups="keep")

@@

```{r}
## Create HHI Level data
### On CSA Level ###
landfill_shrs_csa <- landfill_csa_df %>%
  ungroup() %>%
  group_by(Owner_Code,market,iscsa) %>% # Aggregate on Market level, Owner Entity, Operator level
  summarise(waste_own = sum(`Total of All Wastes (tons/day)`), .groups="keep") %>%
  ungroup() %>%
  group_by(market) %>% # Aggregate on Market
  mutate(total_waste = sum(waste_own),
         waste_shr = 100*(waste_own/total_waste),
         waste_hhi = sum(waste_shr^2))
### On CBSA Level ###
landfill_shrs_cbsa <- landfill_cbsa_df %>%
  ungroup() %>%
  group_by(Owner_Code,market,iscbsa) %>% # Aggregate on Market level, Owner Entity, Operator level
  summarise(waste_own = sum(`Total of All Wastes (tons/day)`), .groups="keep") %>%
  ungroup() %>%
  group_by(market) %>% # Aggregate on Market
  mutate(total_waste = sum(waste_own),
         waste_shr = 100*(waste_own/total_waste),
         waste_hhi = sum(waste_shr^2)) ## Have NA remove option

@@

@@

@@

## Data check if all percentages in a market equal 1
chk_lf_df <- landfill_shrs_csa %>% 
  group_by(market) %>%
  mutate(chk_per=sum(waste_shr)) %>%
  ungroup() %>%
  select(chk_per) %>%
  unique()
```
### Chatanooga County

```{r}
chat_sum <- chat_shrs %>%
  ungroup() %>%
  select(Owner_Code,waste_own,waste_shr,waste_hhi)

@@

```{r}
chat_sum %>%
  select(Owner_Code,waste_own,waste_shr) %>%
  arrange(desc(waste_shr)) %>%
  kbl(col.names=c("Owner Code","Total Waste","Waste Share"),
      digits=0,
      caption = "# Landfills by Markets") %>%
  kable_classic_2(full_width = F)
```

@@

## CBSA level bar chart data
bar_cbsa <- landfill_shrs_cbsa %>% 
  ungroup() %>%
  group_by(market) %>%
  summarise(landfills_cnt = n()) %>%
  ungroup() %>%
  group_by(landfills_cnt) %>%
  summarise(freq=n()) %>%
  mutate(grp_lbls = ifelse(landfills_cnt>=10,"10+",landfills_cnt)) %>%
  ungroup() %>%
  group_by(grp_lbls) %>%
  summarise(freq=sum(freq)) %>%
  mutate(lvl="CBSA")

@@

```{r}
### Check to see if they are the same
chk_vertical <- disposal_df %>%
  filter(Type=="LF"|Family=="LF"|Type=="TS"|Family=="TS") %>%
  mutate(is_municp=grepl("COUNTY|CITY|TOWN|VILLAGE",Owner_Code),
         is_municp_pub=ifelse(Operation=="Public",TRUE,FALSE),
         is_equal=is_municp==is_municp_pub)

landfill_shrs_csa_operation <- landfill_csa_df %>%
  ungroup() %>%
  group_by(Owner_Code,market,iscsa,Operation) %>% # Aggregate on Market level, Owner Entity, Operator level
  summarise(waste_own = sum(`Total of All Wastes (tons/day)`), .groups="keep") %>%
  ungroup() %>%
  group_by(market) %>% # Aggregate on Market
  mutate(total_waste = sum(waste_own),
         waste_shr = 100*(waste_own/total_waste),
         waste_hhi = sum(waste_shr^2))
```

@@

vertical_firms <- landfill_shrs_csa %>%
  mutate(is_municp=grepl("COUNTY|CITY|TOWN|VILLAGE",Owner_Code)) %>%
  group_by(is_municp,market) %>%
  summarise(freq=n()) %>%
  mutate(is_municp= factor(is_municp,levels=c(TRUE,FALSE),labels=c("Vertically Integrated","Non-Vertically Integrated"))) %>%
  pivot_wider(names_from=is_municp, values_from=freq) %>%
  mutate(`Non-Vertically Integrated`=ifelse(is.na(`Non-Vertically Integrated`),0,`Non-Vertically Integrated`),
         `Vertically Integrated`=ifelse(is.na(`Vertically Integrated`),0,`Vertically Integrated`),
         `Non-Vertically Integrated`=factor(ifelse(`Non-Vertically Integrated`>=10,"10+",`Non-Vertically Integrated`),levels=c(seq(0,9),"10+"),labels=paste(c(seq(0,9),"10+"))),
         `Vertically Integrated`=factor(ifelse(`Vertically Integrated`>=10,"10+",`Vertically Integrated`),levels=c(seq(0,9),"10+"),labels=paste(c(seq(0,9),"10+"))))
```
